Analysis Tá de Pé

### 2019-07-31 article/article.Rmd
add pars section 2

One of the key determinants of efficient public service provision is a robust accountability system \citep{besley2003incentives,cameron2004public,o1998horizontal}. In its standard definition, accountability is the process of holding authorities into account for their actions \citep{finer1941administrative,mulgan2000accountability,o1990bureaucratic}. Past studies show that accountability ensures that politicians act on behalf of voters \citep{freire2010ngp, moncrieffe1998reconceptualizing}, reduces the opportunities for rent-seeking and corruption \citep{deininger2005does,wenar2006accountability} and increases the efficiency of the public sector \citep{adsera2003you,bjorkman2009power}. Moreover, greater accountability is associated with higher levels of economic growth, not only because it limits state discretion in the economy, but also because it promotes social cohesion and long-term investments in human capital \citep{benhabib2010economic, cox2018executive, suebvises2018social,ponzetto2018social}.

@@

Here we assess the impact of a technology-driven solution that lowers the costs of evaluating public works and punishing political representatives in Brazil. The _Tá de Pé_ (TDP) mobile phone application allows Brazilian citizens to go to school construction sites, check their status, and request information to public authorities. TDP users can also take pictures from delayed works and submit them to a group of trained engineers, who then evaluate the execution of the project. If the engineers classify the construction as delayed, TDP sends a message to the mayor's office asking for completion estimates and explanations about why the building is unfinished. TDP received the 2016 Google Social Impact grant, earning more than 200 thousand popular votes, and it has been online since April 2017 in 1030 Brazilian municipalities.

Contrary to our expectations, the TDP app has delivered mixed results. Two findings are positive and in line with our theoretical predictions. From August 2017 to February 2019, TDP increased the likelihood of school completion by 6.8 percent (SE = X.X, t-test = XXX). Additionally, we find that pressure over the bureaucracy increased the investments in the schools by 157.5 percent (SE = X.XX, t-test = XXX). In contrast, our results show the app had a null impact on school constructions statistics. Brazilian politicians did not respond to the citizen's requests placed via the application. Therefore, our findings raise questions about the ability of citizens to hold representatives accountable using bottom-up strategies.

# The _Tá de Pé_ App
\label{sec:tdp}

>>>>

One of the key determinants of efficient public service provision is a robust accountability system \citep{besley2003incentives,cameron2004public,o1998horizontal}. In its standard definition, accountability is understood as the process of holding authorities into account for their actions \citep{finer1941administrative,mulgan2000accountability,o1990bureaucratic}. Past studies show that accountability ensures that politicians act on behalf of voters \citep{freire2010ngp, moncrieffe1998reconceptualizing}, reduces the opportunities for rent-seeking and corruption \citep{deininger2005does,wenar2006accountability} and increases the efficiency of the public sector \citep{adsera2003you,bjorkman2009power}. Moreover, greater accountability is associated with higher levels of economic growth, not only because it limits state discretion in the economy, but also because it promotes social cohesion and long-term investments in human capital \citep{benhabib2010economic, cox2018executive, suebvises2018social,ponzetto2018social}.

@@

Here we assess the impact of a technology-driven solution that lowers the costs of evaluating public works and punishing political representatives in Brazil. Developed by Transparência Brasil[^transparencia], the _Tá de Pé_ (TDP) mobile phone application allows Brazilian citizens to know the location of public school construction sites, check their completion status, and anonymously request information from competent authorities. TDP users can also take pictures of delayed works and submit them to a group of trained engineers for examination. If the engineers classify the construction as delayed, TDP sends a message to the mayor's office asking for completion estimates and explanations about the construction status. TDP has been online since April 2017 in 1030 Brazilian municipalities and was the winner of the 2016 Google Social Impact grant with more than 200 thousand popular votes[^google].

[^transparencia]: Transparência Brasil is an non-governmental organisation whose mission is to 'promote transparency and social control of public power' in Brazil. It has been active since April 2000, is non-partisan and receives no public funding. More information is available at <http://transparencia.org.br> (access: July 2019).

[^google]: About 1,000 Brazilian charities participated in the 2016 Google Social Impact Challenge and an independent selection committee selected 10 of them as finalists. Transparência Brasil was then the highest voted NGO by the public. To know more about the contest, please visit <https://impactchallenge.withgoogle.com/brazil2016> (access: July 2019).

We used TDP to conduct three experimental interventions. We tested whether the app has a positive impact on school completion rates, school construction status and complaints to competent authorities. Contrary to our expectations, the TDP app has delivered mixed results. Two of the findings are positive and in line with our theoretical predictions. From August 2017 to February 2019, TDP increased the likelihood of school completion by 6.8 percent (SE = X.X, t-test = XXX). Additionally, we find that pressure over the bureaucracy increased the investments in the schools by 157.5 percent (SE = X.XX, t-test = XXX). In contrast, our results show the app had a null impact on school constructions statistics. Brazilian politicians did not respond to the citizen's requests placed via the application. Therefore, our findings raise questions about the ability of citizens to hold representatives accountable using bottom-up strategies. **Nota**: _Reescrever essa última frase e colocar mais explicações sobre possíveis mecanismos, etc_

# The _Tá de Pé_ Project \label{sec:tdp}

The _Tá de Pé_[^tdp] cell phone application is an initiative carried out by Transparência Brasil to foster bottom-up accountability in the Brazilian public sector. More specifically, the main goal of the TDP project is to improve the responsiveness in education infrastructure public expenditures. As we mentioned above, TDP incentivises citizens to provide up-to-date information about unfinished schools in their neighbourhoods, and these pictures are then assessed by engineers. In the case the construction is behind schedule, citizens are also provided with tools to contact their representatives quickly and anonymously. The app sends a notification to the mayor's office which has 15 days to answer. If they fail answer by the deadline, a notification goes to the federal government, making it harder to the municipality to access federal funds in the future. The motivation for the intervention is that information delivery causes individuals to closely monitor public works. This, in turn, results in better social outcomes as public agents become more responsible to community demands.

[^tdp]: _Tá de Pé_ is an informal Brazilian expression for 'is it done?'. Literally, it means 'standing on its feet' in Portuguese.

We built the app in the first three months of 2017 and tested it in May 2017. The first stable Android version was deployed on Google Play on 14 August 2017. A version for iOS appeared two weeks later. At the same time, a massive Facebook campaign started in order to publicise the app. Facebook is the most widely used social media network in Brazil with around 72 million users in 2017 [@statista2019facebook]. The campaigns were concentrated in October 2017, investing around BRL 14.333,74 (US$ 4.479,29), and attracting 2028 new users in October 2017 only.

\vspace{.2cm}

\begin{figure}
     \begin{minipage}[t]{.3\textwidth}
       \centering
       \includegraphics[scale=0.25]{tdp01.jpg}
       \label{}
     \end{minipage}
     \hfill
     \begin{minipage}[t]{.3\textwidth}
       \centering
       \includegraphics[scale=0.25]{tdp02.jpg}
     \end{minipage}
     \hfill
     \begin{minipage}[t]{.3\textwidth}
       \centering
       \includegraphics[scale=0.25]{tdp03.jpg}
     \end{minipage}
\caption{The \textit{Tá de Pé} mobile phone application. The first image shows that a school construction is delayed by 9 months. The second image shows how citizens can add information to the photos they submit via the app. The last image presents a list of the closest school construction sites.} 
\label{fig:screenshots}
\end{figure}

Transparência Brasil established a partnership with the Brazilian branch of _Engineers without Borders_ (EWB), a non-governmental organisation dedicated to promote international development[^ewb], to provide unbiased evaluations of school completion rates. Their assessment is based on photos and GPS location sent through the app. All data are collected anonymously and neither organisation are able to identify which user uploaded the images. Information on users' alerts are stored on their mobile devices, so citizens can track whether their request has been answered by public authorities.

[^ewb]: Please visit <http://www.ewb-international.org> to know more about Engineers without Borders International and <https://esf.org.br> for information on the Brazilian office.

The TDP has also received feedback from Brazilian computer programmers and policy analists. In 2017 and 2018, Transparência Brasil announced two team programming competitions, called 'Tá de Pé Hackathons', where contributors could fix code bugs and suggest new functionalities to TDP. In total, 75 participants

### 2019-07-31 tdp/analysis3.Rmd
add pars section 2

---
title: "Analysis Tá de Pé"
author: "Umberto Mignozzetti"
date: "7/26/2019"
header-includes:
   - \usepackage[utf8]{inputenc}
   - \usepackage{tipa}
   - \usepackage{caption,tabularx,booktabs,dcolumn}
output: pdf_document
---

```{r, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
####
## Pre-processing
####

@@

# Data block randomization
setwd("~/Dropbox/Academic/Projects/TDP/2_analysis/")
load('./bucket/final_bucket.RData')

>>>>

---
title: "Analysis Tá de Pé"
author: ''
date: "7/26/2019"
output:
  html_document:
    df_print: paged
header-includes:
- \usepackage[utf8]{inputenc}
- \usepackage{tipa}
- \usepackage{caption,tabularx,booktabs,dcolumn}
---

```{r, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
r <- getOption("repos")
r["CRAN"] <- "http://cran.cnr.berkeley.edu/"
options(repos = r)

####
## Pre-processing
####

@@

# Data block randomization
setwd("/Users/danilo/Documents/github/TDPImpactEval/tdp/")
load('final_bucket.RData')
